\documentclass[14pf, a4paper]{article}
\usepackage[english]{babel}
\usepackage{palatino}
\usepackage[utf8]{inputenc}
\usepackage[hidelinks]{hyperref}
\usepackage{nth}

\title{The Apple Never Falls Far From The Tree}
\author{David Kelly}
\date{April 2016}

\begin{document}

\maketitle

\section{Background}
The apple is a fascinating plant. There are presently over 7,500 \textit{cultivars} of apple
known\footnote{\url{https://en.wikipedia.org/wiki/Apple}}. The genome is  incredibly complex, with
approximately 57,000 genes, almost double that of humans. If one were to plant a apple from a Granny
Smith tree, the resulting tree and fruit would be unique. All Granny Smith's are grown from cuttings
from the original tree. This composition tries to reflect some of these characteristics.

The English folk song \textbf{I will give my love an apple}\footnote{\url{http://www.thecanadianencyclopedia.ca/en/article/ill-give-my-love-an-apple-emc/}}, 
dating from \nth{15} century\footnote{here in a performance by Andreas Scholl} is used as the basis of a piece of music concréte which
progressively transforms the source material through a series of mutations, much in the manner that
the apple genome itself morphs from generation to generation\footnote{Lines from Hardy's poem
    'Heredity' come to mind: I am the family face;/ Flesh perishes, I live on,/ Projecting trait
    and trace/ Through time to times anon,/ And leaping from place to place/
Over oblivion.}. Some generations are similar to
their parents, some are far removed. Obvious progenitors for the composition would include Steve
Reich's seminal tape piece \textit{It's Gonna Rain}\footnote{Reich, 1965} which uses found sound and
process-driven composition in a similar fashion. Another source of inspiration was Luciano Berio's
\textit{Sequenza III for Human Voice}\footnote{also, by coincidence, from 1965}. Here, text, or 
rather words, are split into constituent parts, or phonemes, and recombined to create new
groupings.

\section{Musical Form}
The composition loosely follows sonata form. There is an initial introduction and exposition of the
material (an extract from the source recording) which is then sliced into sections. Imagine these sections to be
representative of genetic strands, mobile and malleable. Next, there is a
\textit{pseudo}-development, an attempt to mimic the transition in sonata-form. This uses,
initially, a purely random duplication of some elements, followed by a process controlled by a first-order
Markov chain\footnote{\url{http://explodingart.com/jmusic/jmtutorial/Markov1.html}}. 
This terminates with a brief recapitulation, before the process of mutation starts. A randomized population is tested for
\textit{fitness} against the original seed (that is, the original musical extract)\footnote{The
    method for this is explained more fully in the ipython presentation}. Slowly, the
fragments move further and further from the source material, until only fugitive elements are left,
leading the listener back into the final, and real, recapitulation.

\section{Technical Details}
The program is written in Python3.5, using the pydub library for audio sampling and numpy for fast
mathematical procedures. As each performance will be different, a recording of one possible
performance has been submitted. Running the program again should result in a unique, but genetically
related, composition.
\end{document}
